---
title: "Files loading and description"
author: "Aude Dejonghe and Carlos Biddle"
format: html
---

```{r}
#| message: false
library(here)
here::i_am("R-project-2024_Biddle-Dejonghe.Rproj")
```
# INTRODUCTION

This document will show :
- the sources used for the project with their description 
- the description of the work of joining and cleaning the data
- the explanation of our research question

To begin with, here is the link to the github project : https://github.com/audedjgh/R-project-2024_Biddle-Dejonghe

```{r}
#| message: false
library(vroom)
library(dplyr)
```


# FILES IMPORT

```{r}
#| message: false

library(readr)
library(readxl)

Better_Life_Index_2022 <- read_delim(
  file = "data/Better Life Index 2022.csv", 
  delim = ";", 
  escape_double = FALSE, 
  trim_ws = TRUE,
  show_col_types = FALSE
)

suppressMessages(Education <- read_excel(
  path = "data/Education.xls", 
  sheet = "Data"
))

suppressMessages(GDP <- read_excel(
  path = "data/GDP (current US$).xls", 
  sheet = "Data"
))

suppressMessages(Gas_emissions <- read_excel(
  path = "data/GovReg_GHG proxy_2023.xlsx", 
  sheet = "EEA proxy dataset (plus)"
))

suppressMessages(Health <- read_excel(
  path = "data/Health.xls"
))

suppressMessages(Share_of_renewable_sources <- read_excel(
  path = "data/Share of energy from renewable sources.xlsx", 
  sheet = "EU27"
))
```


## Link to sources

Better Life Index 2022 : https://www.oecd-ilibrary.org/social-issues-migration-health/data/oecd-social-and-welfare-statistics/better-life-index-edition-2022_a4815531-en

Education : https://data.worldbank.org/topic/education

GDP : https://data.worldbank.org/indicator/NY.GDP.MKTP.CD?locations=EU

Gas emissions : https://www.eea.europa.eu/en/datahub/datahubitem-view/f6e68f73-b494-4f8c-8c73-8a153a53f64a

Health : https://data.worldbank.org/topic/health

Share of renewable energy : https://ec.europa.eu/eurostat/web/energy/database/additional-data#Short%20assessment%20of%20renewable%20energy%20sources%20(SHARES)




# DESCRIPTION OF THE SOURCES

## Better Life Index 
Our first dataset includes data for the OECD's Better Life Index in 2022. The OECD Better Life Index Edition 2022 provides data on well-being across OECD member countries and some partners (we'll concentrate on 27 EU countries), covering topics such as housing, income, jobs, education, community, health, environment, safety, life satisfaction, and work-life balance. The data is compiled from national statistical offices, OECD surveys, and other relevant tools, ensuring global comparability of quality-of-life metrics. This dataset is updated annually, though some components, like subjective well-being or community engagement metrics, depend on periodic survey collections. It enables cross-country comparisons and offers insights into policy impacts on citizens' lives.
Since our research is going to compare data over years, this is an example of the better life index, but we'll download and summarise other files like this to have an overview time wise. The data is available since 2013, each year.

|                        |                      |
|------------------------|----------------------|
| Number of observations | `r NROW(Better_Life_Index_2022)` |
| Number of variables    | `r NCOL(Better_Life_Index_2022)` |


## Education
The education dataset from the World Bank database has 165 indicators to choose from, including enrollment rates, literacy levels, teacher-pupil ratios, government spending on education, and learning outcomes. This data is drawn from government ministries, international organizations, and World Bank studies. It is updated annually, though some indicators, like results from international assessments (e.g., PISA, TIMSS), are released every few years. The dataset supports monitoring of global education progress, disparities, and achievements, particularly in relation to Sustainable Development Goal 4 on quality education.
We'll consider all 27 EU countries - into 4 income groups (high income, upper middle income, lower middle income, low income).

|                        |                      |
|------------------------|----------------------|
| Number of observations | `r NROW(Education)` |
| Number of variables    | `r NCOL(Education)` |


## GDP
This dataset provides annual Gross Domestic Product (GDP) figures in current US dollars for countries worldwide, with detailed data on the European Union. The World Bank collects this information from national accounts, statistical agencies, and IMF sources, ensuring consistency across countries. Updates occur annually, with quarterly estimates provided for certain regions. This dataset is vital for evaluating economic performance, identifying growth trends, and comparing the GDP of different regions and countries over time.
It's calculated with official exchange rates, without adjusting for inflation or purchasing power parity, and represents the monetary value of all finished goods and services produced in an economy.

|                        |                      |
|------------------------|----------------------|
| Number of observations | `r NROW(GDP)` |
| Number of variables    | `r NCOL(GDP)` |


## Gas emissions
The European Environment Agency (EEA) Renewable Energy Data tracks the use and contribution of renewable energy in Europe, including electricity, heating, and transport sectors. It estimates the share of renewables in gross final energy consumption and monitors progress toward EU renewable energy targets. This data is collected from national energy statistics and harmonized through Eurostat and other EU reporting mechanisms. It is updated annually, following the SHARES (Short Assessment of Renewable Energy Sources) methodology, ensuring alignment with EU directives. The dataset is a key resource for assessing the EU’s progress toward climate and energy goals.
It offers estimates for total emissions along with industry-specific contributions, for instance, those relating to energy, transport, and agriculture. The data is crucial for policymakers, researchers and environmentalists to study progress in emission reductions.

|                        |                      |
|------------------------|----------------------|
| Number of observations | `r NROW(Gas_emissions)` |
| Number of variables    | `r NCOL(Gas_emissions)` |


## Health
Then, we have the health dataset, supplied by the World Bank.It includes indicators such as life expectancy, healthcare expenditure, immunization rates, and disease prevalence. The data is sourced from national health agencies, WHO publications, and World Bank studies. It is updated annually, with some metrics dependent on periodic demographic and health surveys (e.g., DHS) or WHO global health reports. This dataset enables the analysis of health outcomes, resource allocation, and comparisons of health system performance across countries and regions.

|                        |                      |
|------------------------|----------------------|
| Number of observations | `r NROW(Health)` |
| Number of variables    | `r NCOL(Health)` |


## Share of renewable energy
Finally, the Eurostat Renewable Energy Database (SHARES) focuses on tracking renewable energy trends in the EU, measuring energy production, installed capacity, and the proportion of renewables in the energy mix. It also evaluates progress toward national and EU-wide renewable energy goals. Data is collected from national statistical offices and harmonized through EU legislative frameworks. Updates are conducted annually, using the SHARES methodology, which ensures consistent reporting and alignment with EU policies. This dataset provides detailed insights for policymakers and stakeholders engaged in the EU Green Deal and sustainable energy initiatives.
It provides in-depth statistics on the fraction of renewable energy sources to gross final energy consumption in EU member states. The data covers electricity, heating and cooling, and transport. Key points include national shares of renewable energy, normalised values for hydropower and wind power, and compliance with sustainability criteria.

|                        |                      |
|------------------------|----------------------|
| Number of observations | `r NROW(Share_of_renewable_sources)` |
| Number of variables    | `r NCOL(Share_of_renewable_sources)` |




# MOST IMPORTANT VARIABLES

With the Better Life Index, the value associated to each country is calculated as a weighted score of 11 topics. The topics are housing, income, jobs, community, education, environment, governance, health, life satisfaction, safety, and work-life balance.

As for the education dataset, we can focus on the variable "Educational attainment, at least completed upper secondary, population 25+, total (%) (cumulative)", as it measures the proportion of the adult population with at least a high school diploma, which is a crucial idnicator of human capital development, and helps understand a country's workforce competitiveness.


